---
title:  "Intro to R HW"
author: "Matthew Bentz"
date:   "1/19/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1 & 2

Importing the data from 'baseball.csv' as a data frame and checking for 'Factor' structures.

```{r}
baseball <- read.csv("C:/Users/matth/Downloads/baseball.csv")

is.factor(baseball$bat)
is.factor(baseball$throw)
is.factor(baseball$field)
```

# Exercise 3

Create a subset of pitchers and calculate the mean standard deviation of the ERA.

```{r}
pitchers <- subset(baseball, field == 0)

sd(pitchers$average)
```

# Exercise 4

Create a subset of fielders and calculate the mean standard deviation of the batting average.

```{r}
fielders <- subset(baseball, field == 1)

sd(fielders$average)
```

# Exercise 5

Define BMI and add it to the data frame.

```{r}
bmi <- (baseball$weight * 703) / baseball$height^2

baseball$bmi <- bmi
```

# Exercise 6

Sort the values in increasing order by BMI.

```{r}
baseball <- baseball[order(baseball$bmi, decreasing = FALSE),]

print(baseball)
```

# Exercise 7

Comparison boxplot of the BMI's of pitchers and fielders.

```{r}
boxplot(subset(baseball, field == 1)$bmi, subset(baseball, field == 0)$bmi,
        names=c("Fielders", "Pitchers"))
```

# Exercise 8

Mean/Standard Deviation of the heights, weights, and BMI's of fielders.

```{r}
fielders <- subset(baseball, field == 1)

# --- Mean --- #

## Height
mean(fielders$height)

## Weight
mean(fielders$weight)

## BMI
mean(fielders$bmi)

# --- Standard Deviation --- #

## Height
sd(fielders$height)

## Weight
sd(fielders$weight)

## BMI
sd(fielders$bmi)
```

# Exercise 9

The difference between the avg. ERA of pitchers who's BMI is >= 25 and the avg. ERA of pitchers with BMI < 25, shown as absolute value.

```{r}
pitchers <- subset(baseball, field == 0)

abs(mean(subset(pitchers, bmi >= 25)$average) - mean(subset(pitchers, bmi < 25)$average))
```

# Exercise 10

Create a new data frame that contains overweight players based on BMI.

```{r}
owbb <- subset(baseball, bmi >= 25)

print(owbb)
```

















